Transition path sampling based calculations of free energies of enzymatic reactions: the case of human methionine adenosyl transferase and plasmodium vivax adenosine deaminase

### 2022-05-04 tex/manuscript.tex
made some more corrections

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jpcbfk,manuscript=article,layout=twocolumn]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{chemformula} % Formula subscripts using \ch{}
\usepackage[T1]{fontenc} % Use modern font encodings
\usepackage{graphicx}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{mathptmx}
%\usepackage[scaled=0.92]{helvet}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

@@

Free energy calculations based on the TPS shooting algorithm is 
appealing but not a lot of examples exist in the literature 
which compare the computational calculations with experimental
results for enzymatic reactions.  
In this manuscript we implement and apply the 
algorithm developed by Radhakrishnan et al. \cite{Radhakrishnan04JChemPhys121p2436} 
to calculate the free energies of reactions catalyzed by two systems:
the human MAT2A enzyme and the plasmodium vivax adenosine deaminase 
enzyme. \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
These enzymes were chosen not only because of their importance in various 
biochemical applications but also because of the availability of experimental 
results with which comparisons can be drawn. The manuscript is organized as 
follows: first we discuss the TPS method and the algorithm that we use 
to calculate the free energies briefly, then we show calculations which apply this 
method on the two enzymatic reactions. Finally we compare our results to expeiments
and provide conclusions. 
%------------------------------------------------------------------------------ 
\section{Methods}
%------------------------------------------------------------------------------ 
\subsection{Transition path sampling}
%------------------------------------------------------------------------------ 
Here we briefly describe the equations describing the TPS method for the purpose of
introducing notations as well as for the sake of being complete. 
Consider the molecular dynamics simulation of a molecular system consisting 
of $N$ atoms. Starting from the initial positions and momenta of each atom at time $t_0$,
the time evolution is carried out by finding the position ($\textbf{q}$) and 
momenta ($\textbf{p}$) of each atom at regular intervals of time $t_i = t_0 + i\Delta t$  
($i = 0,1,2,\ldots$). At each of the time slices, the positions and 
momenta can be collectively represented as
the set $z = \{\textbf{q},\textbf{p}\}$. If the MD simulation is 
run for a total 
time of $\mathcal{T}$ the number of time slices is given by 
$L = \mathcal{T}/\Delta t +1$ and for this sequence of 
times the state of the system can be represented as 
\begin{equation}
z(\mathcal{T}) = \{z_0, z_{\Delta t}, z_{2\Delta t},\ldots,z_{\mathcal{T}}\}
\end{equation}
The probability to obtain a particular sequence of states is determined by the initial 
conditions and the type of dynamics used for the time evolution. If the time evolution is 
Markovian, the probability to go from $z_{i\Delta t}$ to $z_{(i+1)\Delta t}$ depends only 
on $z_{i\Delta t}$ and not on conditions before the time $i\Delta t$ the total path probablility can 
be expressed as the product of individual probabilities $p(i\rightarrow j)$ as 
\begin{equation}
\mathcal{P}[z(\mathcal{T})] = \rho(z_0)\Pi_{i=0}^{\mathcal{T}/\Delta t-1} p(z_{i\Delta t}\rightarrow z_{(i+1)\Delta t}),
\end{equation}
where $\rho(z_0)$ is the equilibrium distribution of the initial conditions and for a canonical ensemble this can be 
expressed as 
\begin{equation}
\rho(z_0) = \exp(-\beta H(z_0))/Z
\end{equation}
where 
\begin{equation}
Z = \int dz \exp(-\beta H(z)) 
\end{equation}
is the canonical partition function.

@@

The reaction catalyzed by the human MAT2A enzyme is
between adenosyl $5'$-triphosphate (ATP) and L-methionine (MET) molecules resulting in the formation of
S-adenosyl-L-methionine (SAM) which is depicted in Fig. \ref{fig:mat2a-reaction}.
SAM is an essential metabolite which is distributed to almost all 
body tissues and fluids, is an universal methyl donor and is of
fundamental importance to the metabolism of compounds such as hormones, neureotransmitters, proteins, 
and nucleic acids. \cite{Friedel89Drugs38p389} This reaction was experimentally characterized by 
Firestone et al. \cite{Firestone17JAmChemSoc139p13754} to follow a S$_{\text{N}}2$ transition 
state. The key bond parameters for this reaction consists of the distance between the
nucleophilic sulfur atom of MET and the $5'-C$ atom of the ATP (denoted as $dis(SC)$), and the distance between 
oxygen atom of the phosphate group and the $5'-C$ atom of ATP ($dis(OC)$).
For the TPS calculations we define the reactant state to have $dis(OC)$ to be 
less than 1.7 {\AA} and the product state to have the $dis(SC)$ less than 1.7 {\AA}.

@@

$pv$ADA catalyzes the reaction between adenosine and hydroxide ion 
resulting in the formation of inosine and ammonia as shown in
Fig. \ref{fig:ada-reaction}.
$pv$ADA is an essential enzyme in the purine salvage pathway and 
inhibiting this enzyme is a promising approach to treat malaria. \cite{Madrid08JBiolChem283p35899}  
This reaction proceeds with the nucleophilic O atom (O$^{\text{hydroxyl}}$) of the hydroxide anion
attacking the electrophilic C6 atom of the adenosyl ring leading to the formation of a Meisenheimer 
complex while the N1 atom on the adenosyl ring starts to get 
protonated by the Glu229 residue of $pv$ADA. At the transition state, experimental findings 
suggest that the dissociation of the C6$-$N6 bond as well as the protonation of the N1 
atom are almost complete. \cite{Luo07JAmChemSoc129p8008} The key bond parameters are 
the distance between the approaching hydroxyl oxygen and the C6 atom (denoted as $dis(OC)$) and the 
dissociating C6$-$N6 ($dis(NC)$) bond distance. For the TPS simulaions we define the reactant state to 
have $dis(NC)$ being less than $1.7$ {\AA} and the product state to have the 
$dis(OC)$ less than $1.7$ {\AA}.

@@

A biased initial reactive trajectory is obtained by using harmonic restraint forces 
of 1500 to break the $O-C$ bond and form the $S-C$ bond for the MAT2A enzyme and 
to break the $N-C$ bond and form the $O-C$ bond for the case of the $pv$ADA enzyme. 
Using the biased trajectory as the starting point and after removing the restraint 
forces, the shooting algorithm was used to generate
new trajectories within which first a random time slice is chosen from the reactive 
trajectory and perturbing the momenta of the system. 
The time step that we used for all our calculations was $1\;fs$ and we propagate the 
dynamics for a total of 250 steps forward and 250 steps backward which amounts to a 
total of $0.5\;ps$. 
A total of 150 and 100 reactive trajectories were generated for the MAT2A and $pv$ADA
enzymes, respectively. 
%------------------------------------------------------------------------------ 
\subsection{Free energy calculations}
%------------------------------------------------------------------------------ 
To obtain the free energy profile along the order parameter, we use the method 
discussed in sec. \ref{ssec:algorithm}. The order paratemer for the MAT2A system 
is defined to be $dis(OC)-dis(SC)$ and that for the $pv$ADA is defined as 
$dis(NC)-dis(OC)$. For the MAT2A system, configurations are sampled within 
the range $[-2.0,2.0]$ and this range is divided into 14 overlapping windows 
($0.09$ {\AA} overlap on either side). 
Configurations are sampled for the $pv$ADA system within the interval $[-3.0, 3.0]$ {\AA}
of the order parameter value. 
This interval is divided into 20 overlapping windows ($0.045$ {\AA} between 
neighboring windows). A reactive trajectory from the TPS ensemble is chosen to be the guide trajectory 
for sampling within each of the windows. To sample in a particular window $w$, a 
frame from this reactive trajectory is chosen such that the value of its 
order parameter falls within $w$. Using the shooting algorithm
the momenta are perturbed and dynamics is run for 10 steps forwards and backwards with a 
$1\;fs$ time step to obtain a trajectory with a time period of $20\;fs$.    
Within each window 3000 TPS trajectories are collected. The 
probability distribution function for the order parameter within window $w$ 
is then calculated as normalized histograms which represent the populations 
of the configurations within the window at equilibrium. 
Then we calculate the free energy profile in a similar fashion to the WHAM 
procedure \cite{Kumar92JComputChem13p1011} which is generally used to obtain the 
potential of mean force from umbrella sampling calculations. 
%------------------------------------------------------------------------------ 
\section{Results and discussion}
%------------------------------------------------------------------------------ 
\subsection{Human Methionine Adenosyltransferase 2A}
The equilibrated structure of the MAT2A enzyme in complex with 
ATP, MET and two Mg$^{2+}$ ions is shown in Fig. \ref{fig:mat2a-equil}.
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.2]{figures/mat2a-equil.png}
\caption{Equilibriated structure of the MAT2A enzyme in complex with ATP and 
MET along with 2 Mg$^{2+}$ ions.}
\label{fig:mat2a-equil}
\end{figure}
In the binding pocket the adenosyl moiety and the Phe250 and Ile117 residues 
are found to have $\pi-\pi$ interactions stabilizing the molecule. The triphosphate 
moiety forms hydrogen bonds with ... residues.

@@

\begin{figure}
\includegraphics[scale=1.0]{figures/mat2a-fenergy.pdf}
\caption{The free energy as a function of the order parameter.}
\end{figure}
From our TPS based calculations we calculated the free energy barrier to be 
$16\;\text{Kcal mol}^{-1}$.
%------------------------------------------------------------------------------ 
\subsection{Adenosine deaminase reaction coordinates and free energy}
%------------------------------------------------------------------------------ 
The equilibrated structure for the ADA enzyme complexed with adenosine, 
OH$^{-}$ anion and Zn$^{2+}$ cation is shown in Fig. \ref{fig:ada-equil}.
The Zn$^{2+}$ cation interacts with the His42, His44 and His226 residues
and stabilizes the OH$^{-}$ anion in the pocket.

Glu229 is part of the 
QM region and it acts as a proton donor to the adenosine molecule.

Experimentally KIE studies 
from Schramm and coworkers \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
predict the transition state structures for the plasmodium falciparum, human and bovine 
ADA enzymes. They reported inhibition studies on the plasmodium vivax ADA by using 
MT coformycin as the inhibitor molecule.

The formation of Meissenheimer complex is part of the 
transition state as well as the partial protonation of the 
N1 atom on the adenosyl moity as observed from experiments. 
The reaction free energy is supposed to have two peaks with 
an intermediate state and we observe the same from our calculations. 
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.2]{./figures/ada-equil.png}
\caption{Equilibriated structure of ADA protein}
\label{fig:ada-equil}
\end{figure}

\begin{figure}
\includegraphics[scale=1.0]{figures/ada-diff60.pdf}
\caption{A reactive trajectory}
\label{fig:ada-reactive-traj}
\end{figure}

\begin{figure}
\centering
\includegraphics[scale=0.12]{blender-images/ada/new-ada-trans.png}
\label{fig:ada-trans}
\caption{The transition state structure of the reaction catalyzed by the ADA enzyme.
The formation of the Meissenheimer complex is the rate limiting step and the N-C distance is 
1.60 {\AA} and the O-C distance is 1.48 {\AA}.}
\end{figure}

>>>>

\pdfoptionpdfminorversion=4

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jpcbfk,manuscript=article,layout=traditional]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{chemformula} % Formula subscripts using \ch{}
\usepackage[T1]{fontenc} % Use modern font encodings
\usepackage{graphicx}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{mathptmx}
\usepackage[scaled=0.92]{helvet}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

@@

Free energy calculations based on the TPS shooting algorithm is 
appealing but rare in the literature. Applications of this 
method to study enzymatic reactions and comparisons of the calculated 
free energies to experimental measurements can shine light on the 
accuracy of this approach compared to the standard methods used for 
free energy calculations.    
In this manuscript we implement and apply the 
algorithm developed by Radhakrishnan et al. \cite{Radhakrishnan04JChemPhys121p2436} 
to calculate the free energies of reactions catalyzed by two systems:
the human MAT2A enzyme and the plasmodium vivax adenosine deaminase 
enzyme. \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
Human MAT2A catalyzes the formation of S-adenosyl L-methionine (SAM).
SAM is an essential metabolite which is 
distributed to almost all body tissues and fluids, is an universal 
methyl donor and is of fundamental importance to the metabolism of 
compounds such as hormones, neureotransmitters, proteins, 
and nucleic acids. \cite{Friedel89Drugs38p389}
plasmodium vivax is a parasite that is responsible for the most number 
of cases of malaria globally. Targeting proteins that are essential to
the metabolism of this bacteria such as the adenosine deaminase enzyme 
is a good anti-malarial drug development strategy. \cite{Madrid08JBiolChem283p35899} 
These enzymes were chosen not only because of their importance in 
biochemical applications but also because of the availability of experimental 
results with which comparisons can be drawn. This manuscript is organized as 
follows: first we discuss the TPS method and the algorithm that we use 
to calculate the free energies briefly, then we show calculations which apply this 
method on the two enzymatic reactions. Finally we compare our results to expeiments
and provide conclusions. 
\section{Methods}
\subsection{Transition path sampling}
Here we briefly describe the equations describing the TPS method for the purpose of
introducing notations as well as for the sake of being complete. 
Consider the molecular dynamics simulation of a molecular system consisting 
of $N$ atoms. Starting from the initial positions and momenta of each atom at time $t_0$,
the time evolution is carried out by finding the position ($\textbf{q}$) and 
momenta ($\textbf{p}$) of each atom at regular intervals of time $t_i = t_0 + i\Delta t$  
($i = 0,1,2,\ldots$). At each of the time slices, the positions and 
momenta can be collectively represented as
the set $z = \{\textbf{q},\textbf{p}\}$. If the MD simulation is 
run for a total 
time of $\mathcal{T}$ the number of time slices is given by 
$L = \mathcal{T}/\Delta t +1$ and for this sequence of 
times the state of the system can be represented as 
z(\mathcal{T}) = \{z_0, z_{\Delta t}, z_{2\Delta t},\ldots,z_{\mathcal{T}}\}
The probability to obtain a particular sequence of states is determined by the initial 
conditions and the type of dynamics used for the time evolution. If the time evolution is 
Markovian, the probability to go from $z_{i\Delta t}$ to $z_{(i+1)\Delta t}$ depends only 
on $z_{i\Delta t}$ and not on conditions before the time $i\Delta t$ the total path probablility can 
be expressed as the product of individual probabilities $p(i\rightarrow j)$ as 
\mathcal{P}[z(\mathcal{T})] = \rho(z_0)\Pi_{i=0}^{\mathcal{T}/\Delta t-1} p(z_{i\Delta t}\rightarrow z_{(i+1)\Delta t}),
where $\rho(z_0)$ is the equilibrium distribution of the initial conditions and for a canonical ensemble this can be 
expressed as 
\rho(z_0) = \exp(-\beta H(z_0))/Z
where 
Z = \int dz \exp(-\beta H(z)) 
is the canonical partition function.

@@

The reaction catalyzed by the human MAT2A enzyme is
between adenosyl $5'$-triphosphate (ATP) and L-methionine (MET) molecules resulting in the formation of
S-adenosyl-L-methionine (SAM) which is depicted in Fig. \ref{fig:mat2a-reaction}.
 This reaction was experimentally characterized by 
Firestone et al. \cite{Firestone17JAmChemSoc139p13754} to follow a S$_{\text{N}}2$ transition 
state. The key bond parameters for this reaction consists of the distance between the
nucleophilic sulfur atom of MET and the $5'-C$ atom of the ATP (denoted as $dis(SC)$), and the distance between 
oxygen atom of the phosphate group and the $5'-C$ atom of ATP ($dis(OC)$).
For the TPS calculations we define the reactant state to have $dis(OC)$ to be 
less than 1.7 {\AA} and the product state to have the $dis(SC)$ less than 1.7 {\AA}.

@@

$pv$ADA catalyzes the reaction between adenosine and hydroxide ion 
resulting in the formation of inosine and ammonia as shown in
Fig. \ref{fig:ada-reaction}.
$pv$ADA is an essential enzyme in the purine salvage pathway and 
inhibiting this enzyme is a promising approach to treat malaria.   
This reaction proceeds with the nucleophilic O atom (O$^{\text{hydroxyl}}$) of the hydroxide anion
attacking the electrophilic C6 atom of the adenosyl ring leading to the formation of a Meisenheimer 
complex while the N1 atom on the adenosyl ring starts to get 
protonated by the Glu229 residue of $pv$ADA. At the transition state, experimental findings 
suggest that the dissociation of the C6$-$N6 bond as well as the protonation of the N1 
atom are almost complete. \cite{Luo07JAmChemSoc129p8008} The key bond parameters are 
the distance between the approaching hydroxyl oxygen and the C6 atom (denoted as $dis(OC)$) and the 
dissociating C6$-$N6 ($dis(NC)$) bond distance. For the TPS simulaions we define the reactant state to 
have $dis(NC)$ being less than $1.7$ {\AA} and the product state to have the 
$dis(OC)$ less than $1.7$ {\AA}.

@@

A biased initial reactive trajectory is obtained by using harmonic restraint forces 
of 1500 to break the $O-C$ bond and form the $S-C$ bond for the MAT2A enzyme and 
to break the $N-C$ bond and form the $O-C$ bond for the case of the $pv$ADA enzyme. 
Using the biased trajectory as the starting point and after removing the restraint 
forces, the shooting algorithm was used to generate
new trajectories. The well known shooting algorithm begins with the choice of a 
random time slice from the biased reactive trajectory and perturbing the momenta of the system.
Then the dynamics is propagated for 250 steps forawards and backwards using $1\;fs$ time steps
resulting in a trajectory with a time period of $0.5\;ps$.
A total of 150 and 130 reactive trajectories were generated for the MAT2A and $pv$ADA
enzymes, respectively. 
%------------------------------------------------------------------------------ 
\subsection{Free energy calculations}
%------------------------------------------------------------------------------ 
To obtain the free energy profile along the order parameter, we use the method 
discussed in sec. \ref{ssec:algorithm}. The order paratemer for the MAT2A system 
is defined to be $dis(OC)-dis(SC)$ and that for the $pv$ADA is defined as 
$dis(NC)-dis(OC)$. For the MAT2A system, configurations are sampled within 
the range $[-2.0,2.0]$ and this range is divided into 14 overlapping windows 
($0.09$ {\AA} overlap on either side). 
Configurations are sampled for the $pv$ADA system within the interval $[-3.0, 3.0]$ {\AA}
of the order parameter value. 
This interval is divided into 20 overlapping windows ($0.045$ {\AA} between 
neighboring windows). A reactive trajectory from the TPS ensemble is chosen to be the guide trajectory 
for sampling within each of the windows. To sample in a particular window $w$, a 
frame from this reactive trajectory is chosen such that the value of its 
order parameter falls within $w$. Using the shooting algorithm
the momenta are perturbed and dynamics is run for 10 steps forwards and backwards with a 
$1\;fs$ time step to obtain a trajectory with a time period of $20\;fs$.    
Within each window 3000 TPS trajectories are collected. The 
probability distribution function for the order parameter within window $w$ 
is then calculated as normalized histograms which represent the populations 
of the configurations within the window at equilibrium. 
Then we calculate the free energy profile in a similar fashion to the WHAM 
procedure \cite{Kumar92JComputChem13p1011} which is generally used to obtain the 
potential of mean force from umbrella sampling calculations. 
%------------------------------------------------------------------------------ 
\section{Results and discussion}
%------------------------------------------------------------------------------ 
\subsection{Human Methionine Adenosyltransferase 2A}
The equilibrated structure of the MAT2A enzyme in complex with 
ATP, MET and two Mg$^{2+}$ ions is shown in Fig. \ref{fig:mat2a-equil}.
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.2]{figures/mat2a-equil.png}
\caption{Equilibriated structure of the MAT2A enzyme in complex with ATP and 
MET along with 2 Mg$^{2+}$ ions.}
\label{fig:mat2a-equil}
\end{figure}
In the binding pocket the adenosyl moiety and the Phe250 and Ile117 residues 
are found to have $\pi-\pi$ interactions stabilizing the molecule. The triphosphate 
moiety forms hydrogen bonds with ... residues.

@@

\begin{figure}
\includegraphics[scale=1.0]{figures/mat2a-fenergy.pdf}
\caption{The free energy as a function of the order parameter.}
\end{figure}
From our TPS based calculations we calculated the free energy barrier to be 
$16\;\text{Kcal mol}^{-1}$.
%------------------------------------------------------------------------------ 
\subsection{Adenosine deaminase reaction coordinates and free energy}
%------------------------------------------------------------------------------ 
The equilibrated structure for the ADA enzyme complexed with adenosine, 
OH$^{-}$ anion and Zn$^{2+}$ cation is shown in Fig. \ref{fig:ada-equil}.
The Zn$^{2+}$ cation interacts with the His42, His44 and His226 residues
and stabilizes the OH$^{-}$ anion in the pocket. The N1 atom on the adensoyl 
molecule hydrogen bonds with the carboxylic acid hydrogen atom of the Glu229 
residue.  
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.2]{./figures/ada-equil.png}
\caption{Equilibriated structure of ADA protein}
\label{fig:ada-equil}
\end{figure}
The bond forming ($dis(OC)$) and bond breaking ($dis(NC)$) distances 
for a reactive trajectory from the TPS ensemble is depicted 
in Fig. \ref{fig:ada-reactive-traj}.
\begin{figure}[ht!]
\includegraphics[scale=1.0]{figures/ada-diff60.pdf}
\caption{A reactive trajectory}
\label{fig:ada-reactive-traj}
\end{figure}
Committor analysis is performed to determine the transition state of this reaction
and the structure obtained from the calculations is shown in Fig. \ref{fig:ada-trans}.
At the transition state the N$1-$H bond distance is $1.1\;${\AA}, suggesting that the 
Glu229 residue has completed the proton donation to the adenosine molecule. This protonation
then triggers a base rotation which can be observed    
\begin{figure}
\centering
\includegraphics[scale=0.12]{blender-images/ada/new-ada-trans.png}
\label{fig:ada-trans}
\caption{The transition state structure of the reaction catalyzed by the ADA enzyme.
The formation of the Meissenheimer complex is the rate limiting step and the N-C distance is 
1.60 {\AA} and the O-C distance is 1.48 {\AA}.}
\end{figure}
Experimentally KIE studies 
from Schramm and coworkers \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
predict the transition state structures for the plasmodium falciparum, human and bovine 
ADA enzymes. They reported inhibition studies on the plasmodium vivax ADA by using 
MT coformycin as the inhibitor molecule.

The formation of Meissenheimer complex is part of the 
transition state as well as the partial protonation of the 
N1 atom on the adenosyl moity as observed from experiments. 
The reaction free energy is supposed to have two peaks with 
an intermediate state and we observe the same from our calculations.
